Fig1

### 2020-05-15 analysis/03_datawrangling.Rmd
now with files saved for futuer analysis
now with suppl table and figs

candidategenes <- parentalcaregenes %>% distinct(gene) %>% pull(gene)
# note: I can't find these genes in dataset: NOS1 or OXTR or FOX1B or HTR5A
```

@@

## All DEGs (but currently only char DEGs)

@@

# select only characterization plots
allDEG$tissue <- factor(allDEG$tissue , levels = tissuelevel)
allDEG$comparison <- factor(allDEG$comparison , levels = comparisonlevelschar)
allDEG$direction <- factor(allDEG$direction, levels = charlevels)

allDEG <- allDEG %>% drop_na()
```

@@

```{r write}
write.csv(allDEG, "../results/03_allDEGs.csv")
write.csv(candidatevsd, "../results/03_candidatevsd.csv")
write.csv(parentalcaregenes, "../metadata/03_parentalcaregenes.csv")
```

>>>>

# note: I can't find these genes in dataset: NOS1 or OXTR or FOX1B or HTR5A
```

@@

## All DEGs

@@

unique(allDEG$comparison)
```

@@

```{r write}
write.csv(allDEG, "../results/03_allDEG.csv", row.names = F)
write.csv(candidatevsd, "../results/03_candidatevsd.csv")
write.csv(parentalcaregenes, "../metadata/03_parentalcaregenes.csv")
```

### 2020-05-15 analysis/figure-1.Rmd
now with suppl table and figs

```{r}
allDEG <- read_csv("../results/03_allDEGs.csv") %>%
  mutate(comparison = factor(comparison, levels = comparisonlevelschar))
str(allDEG)
unique(allDEG$comparison)
```

@@

```{r DEGs}

@@

## make figure

@@

d1 <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) + 
  labs(x = NULL, subtitle = "hypothalamus")  + 
  facet_wrap(~sex, scales = "free")
d3 <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors ) + 
  labs(x = NULL, subtitle = "pituitary") + 
  facet_wrap(~sex, scales = "free") +
  theme(strip.text = element_blank())
d5 <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors ) + 
   facet_wrap(~sex, scales = "free") +
  theme(strip.text = element_blank()) +
  labs(x = "tSNE1 \n \n DEGs = differentially expressed genes \n + LFC = positive log fold change", 
       subtitle = "gonads")
d2 <- makebargraph("hypothalamus","DEGs w/ + LFC", 0, 1800) + 
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank())  
d4 <- makebargraph("pituitary","DEGs w/ + LFC", 0, 1800)  +  
  theme(axis.text.x = element_blank(), 
        axis.title.x = element_blank(), 
        strip.text.x = element_blank())   
d6 <- makebargraph("gonad","DEGs w/ + LFC", 0, 1800) +  
  theme(strip.text.x = element_blank()) +
  scale_x_discrete(labels = comparisonlabelschar)+
  labs(subtitle = " ")
d <- plot_grid(d1,d2,d3,d4,d5,d6, ncol = 2, rel_heights = c(1,0.9,1.3), rel_widths = c(1,2),
                  labels = c("D"), label_size = 8)

@@

# supple table 1 of all but control-bldg DEGs

```{r}
suppletable1 <- allDEG %>%
  #filter(comparison != "control_bldg") %>%
  group_by(sex, tissue, comparison) %>%
  arrange( tissue, sex, direction, gene)
head(suppletable1)

write_csv(suppletable1, "../results/suppletable1.csv")

# save file for musical genes https://raynamharris.shinyapps.io/musicalgenes/
#write.csv(allDEG, "../../musicalgenes/data/allDEG.csv")

>>>>

```{r suppltables}
allDEG <- read_csv("../results/03_allDEG.csv") %>%
  mutate(tissue = factor(tissue, levels = tissuelevel),
         direction = factor(direction, levels = alllevels))

# for fig 1
DEGchar <- allDEG %>% 
  filter(comparison %in% comparisonlevelschar) %>%
  mutate(comparison = factor(comparison, levels = comparisonlevelschar))

# for supple fig 1
DEGcontrol <- allDEG %>% 
  filter(grepl("control", comparison),
         !grepl("m.|early|extend|prolong", comparison))  %>%
  mutate(comparison = factor(comparison, levels = comparisonlevelscontrol))

DEGbldg <- allDEG %>% 
  filter(grepl("bldg", comparison),
         !grepl("m.|early|extend|prolong", comparison))  %>%
  mutate(comparison = factor(comparison, levels = comparisonlevelsbldg))

suppltable1 <- DEGchar %>%
  arrange( tissue, sex, direction, gene)
head(suppltable1)
write_csv(suppltable1, "../results/suppltable1.csv")

suppltable2 <- rbind(DEGcontrol, DEGbldg)  %>%
  arrange( tissue, sex, direction, gene)
head(suppltable2)
write_csv(suppltable2, "../results/suppltable1.csv")

@@

```{r DEGs, message=F}

@@

## Figure 1

@@

d1 <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) + 
  labs(x = NULL, subtitle = "hypothalamus")  + 
  facet_wrap(~sex, scales = "free")
d3 <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors ) + 
  labs(x = NULL, subtitle = "pituitary") + 
  facet_wrap(~sex, scales = "free") +
  theme(strip.text = element_blank())
d5 <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors ) + 
   facet_wrap(~sex, scales = "free") +
  theme(strip.text = element_blank()) +
  labs(x = "tSNE1 \n \n DEGs = differentially expressed genes \n + LFC = positive log fold change", 
       subtitle = "gonads")
d2 <- makebargraph(DEGchar, "hypothalamus","DEGs w/ + LFC", 0, 1800) + 
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank())  
d4 <- makebargraph(DEGchar, "pituitary","DEGs w/ + LFC", 0, 1800)  +  
  theme(axis.text.x = element_blank(), 
        axis.title.x = element_blank(), 
        strip.text.x = element_blank())   
d6 <- makebargraph(DEGchar, "gonad","DEGs w/ + LFC", 0, 1800) +  
  theme(strip.text.x = element_blank()) +
  scale_x_discrete(labels = comparisonlabelschar)
d <- plot_grid(d1,d2,d3,d4,d5,d6, ncol = 2, rel_heights = c(1,0.9,1.3), rel_widths = c(1,2),
                  labels = c("D"), label_size = 8)

@@

## suppl fig 1 controls versus bldg

```{r supplfig-1, fig.width=5, fig.height=5}
s1a <- makebargraphsuppl(DEGcontrol, "hypothalamus","DEGs w/ + LFC", 0, 5000) + 
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank()) + 
  labs(subtitle = "hypothalamus") 
s1c <- makebargraphsuppl(DEGcontrol, "pituitary","DEGs w/ + LFC", 0, 5000)  +  
  theme(axis.text.x = element_blank(), 
        axis.title.x = element_blank(), 
        strip.text.x = element_blank())    + 
  labs(subtitle = "pituitary") 
s1e <- makebargraphsuppl(DEGcontrol, "gonad","DEGs w/ + LFC", 0, 5000) +  
  theme(strip.text.x = element_blank()) +
  scale_x_discrete(labels = comparisonlevelscontrol) +
  labs(subtitle = "gonads")

s1b <- makebargraphsuppl(DEGbldg, "hypothalamus", NULL, 0, 1000) + 
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank())  
s1d <- makebargraphsuppl(DEGbldg, "pituitary",NULL, 0, 1000)  +  
  theme(axis.text.x = element_blank(), 
        axis.title.x = element_blank(), 
        strip.text.x = element_blank())   
s1f <- makebargraphsuppl(DEGbldg, "gonad",NULL, 0, 1000) +  
  theme(strip.text.x = element_blank()) +
  scale_x_discrete(labels = comparisonlevelsbldg)

s1 <- plot_grid(s1a,s1b,s1c,s1d,s1e,s1f, ncol = 2, rel_heights = c(1,0.9,1.3))

plot_grid(a, s1, ncol = 1, rel_heights = c(0.25,1))
```

```{r eval = F}
# save file for musical genes https://raynamharris.shinyapps.io/musicalgenes/
#write.csv(allDEG, "../../musicalgenes/data/allDEG.csv")
